---
title: "Stats-Project"
author: "Christopher Wallis"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE}

library(ggplot2)
library(dplyr)
library(knitr)

```


<h3><strong> Dataset </strong></h3>

The following is a data set gathered using environmental tracking websites and contains data on the the meteorologic data in the state of Minnesota as well as the CO2 levels in the atmosphere during the corresponding month and year. 

<u> Year- </u>
<br>
&nbsp; This is the year for the sample.

<u> Month- </u>
<br>
&nbsp; This is the month for the sample.

<u> Precipitation- </u>
<br>
&nbsp; This is the amount of rainfall in inches.

<u> Average Temp- </u>
<br>
&nbsp; This is the average temp during the month for the corresponding sample. This is measured in Celsius.

<u> CO2- </u>
<br>
&nbsp; This is the average CO2 concentration for the corresponding time sample. This is measured in ppm (parts per million).

<h3><strong> Seeting Up The Dataset </strong></h3>

```{r}

environmentalData <- read.csv("StatsProject-Data.csv")

head(environmentalData)

```

```{r}

environmentalDataYearly <- environmentalData %>%
  group_by(Year) %>%
  summarize(Avg_Prec = mean(Precipitation), Avg_Temp = mean(Average.Temp), Avg_CO2 = mean(CO2), digits = 2)

head(environmentalDataYearly)

```



